---
title: "p8105_hw1_ejc2222"
output: github_document
date: "2026-09-27"
---

```{r setup, message = FALSE}
library(tidyverse)
```

# Problem 1

```{r}
data("penguins", package = "palmerpenguins")
```

The penguins dataset contains body measurements for adult penguins collected from 2007 to 2009 near Palmer Station, Antarctica. It covers three species (Adelie, Chinstrap, and Gentoo) found on three islands (Biscoe, Dream, and Torgersen). The key variables include: species, island, sex, and year, along with four physical measurements of bill length and bill depth (mm), flipper length (mm), and body mass (g). The dataset has 344 rows, one per penguin, and 8 columns. The mean flipper length is 200.9 mm.

```{r}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point()

ggsave("penguin_scatterplot.png")
```

A positive relationship shows that penguins with longer bills tend to have longer flippers. The species form different clusters, with Adelie penguins having the shortest bills and flippers, Chinstraps having long bills but flipper lengths similar to Adelies, and Gentoos having the longest flippers of the three.

# Problem 2

This data frame contains a random normal sample, a logical vector, a character vector, and a factor vector.

```{r}
example_df = tibble(
  norm_samp = rnorm(10),
  norm_samp_pos = norm_samp > 0,
  vec_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  vec_factor = factor(c("low", "low", "mid", "mid", "mid",
                        "high", "high", "low", "mid", "high"))
)
```

Taking the mean of each variable.

```{r}
mean(pull(example_df, norm_samp))
mean(pull(example_df, norm_samp_pos))
mean(pull(example_df, vec_char))
mean(pull(example_df, vec_factor))
```

Taking the mean works for the numeric and logical variables, but returns NA with a warning for the character and factor variables.

Convert the logical, character, and factor variables to numeric.

```{r, eval = FALSE}
as.numeric(pull(example_df, norm_samp_pos))
as.numeric(pull(example_df, vec_char))
as.numeric(pull(example_df, vec_factor))
```

Converting the logical variable to numeric turns TRUE into 1 and FALSE into 0, which is why its mean worked. It gives the proportion of sample values greater than 0. The character variable becomes all NAs because letters can't be converted to numbers, matching the failed mean. The factor variable converts to its underlying integer codes (1, 2, 3 for the three levels), but since those codes are just category labels, mean() doesn't treat factors as numeric, which is why that mean also returned NA.